---
title: "A handfull of pixels"
author: "Koen Hufkens"
output: 
  ioslides_presentation:
    css: styles.css
    incremental: false
    widescreen: true
    smaller: false
    logo: "geco_logo_small.png"
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## A handfull of pixels

- Manageable
- Fast proto-typing
- Site based validation data

## Finding & using data {.emphasized}

- Identify how to **get the data**
- Understand the format (extension)
- **Confirm the format**
  - meta-data
  - headers

## Getting data {.emphasized}

- Direct downloads
  - manually
  - not always structured
- Application Programming Interfaces (APIs)
  - automatic

## Application Programming Interfaces (APIs) {.emphasized}

Getting data from only resources (easily)

```{r eval=FALSE}
# load the library
library("MODISTools")

# list all available products
products <- MODISTools::mt_products()
```

## Spatial processing in R {.emphasized}

- Focus on two main tools
- raster data `terra`
- vector data `sf`

Only raster data processing will be discussed in depth.

## Spatial processing in R: package setup {.emphasized}

```{r eval = FALSE}
# Installing the required packages
# (additional software modules)
install.packages("terra")
install.packages("sf")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("MODISTools")
install.packages("here")
```

## Spatial processing in R: project setup {.emphasized}

```{r eval = FALSE}
# Download the project zip file
https://github.com/geco-bern/R_proj_template/archive/refs/heads/main.zip
```

- Download and unzip the file
- Rename the project file (.Rproj)
- Open the Rproj file (opens RStudio)
- Place / save data in `data`
- Place R markdown scripts in `vignettes`

## Spatial processing in R: the basics

```{r eval=TRUE, message=FALSE, warning=FALSE}
# load the library
library("terra")
f <- system.file("ex/elev.tif", package="terra")
elev <- terra::rast(f)

# print the output
print(elev)
```